---
title: "Statistical Inference Project 1"
author: "~\n~"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 6
date: "Friday, December 19, 2014"
---

## Abstract
* In this analysis, the distribution of averages of 40 exponential-distributed random variables is investegated by simulation approach.

* The following questions are answered:
    + Show where the distribution is centered at and compare it to the theoretical center of the distribution.
    + Show how variable it is and compare it to the theoretical variance of the distribution.
    + Show that the distribution is approximately normal.

## Data Analysis

### Random Variables Generation

1. Setting the parameters of the exponential distribution.
```{r, echo=TRUE, eval=TRUE}
library(ggplot2)
library(grid)
# Setting parameters
lambda <- 0.2
# Average number
n <- 40
# Number of simulation
N.sim <- 1000
```

2. Random variables generation.
```{r, echo=TRUE, eval=TRUE}
# Set seed for ruproducible property
set.seed(12345)
# Simulate the average of 40 exp-dist random variables
Sample.sim <- sapply(1:N.sim, function(x) mean(rexp(n,lambda)))
```

### Compare the Center of Theoretical Distrubution and Sample Distribution

```{r, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
# Construct the data frame for Center of Theoretical and sample distribution
Med <- data.frame(Median=c(1/lambda, median(Sample.sim)), 
                  Type=c("Theoretical","Simulation"))
# Visualize the result by ggplot
qplot(x=Sample.sim,geom="histogram") + 
  geom_histogram(fill="skyblue3",colour="darkgrey") + 
  geom_vline(aes(xintercept=Median,colour=Type),data=Med,show_guide = T,size=0.8) +
  scale_colour_manual(values=c("red","black")) + 
  theme(legend.key.width=unit(0.5, "cm"),legend.key.size=unit(1, "cm"),
        legend.position=c(0.8,0.8),legend.background=NULL)
```

### Variance Comparison

The theoretical variance is 
$$var=\frac{1}{40 * lambda^2}=0.625$$. 

Now we compute the sample variance
```{r, echo=TRUE, eval=TRUE}
var(Sample.sim)
```

### The Sample Distribution is Approximately Normal

We can show that the distribution of averages of 40 exponential-distributed random variables is apporximately normal by QQ-plot.
```{r, echo=TRUE,eval=TRUE}
qplot(sample=Sample.sim,stat="qq",distribution = qnorm) + 
  labs(title="Quantile-Quantile Plot: Sample v.s. Normal  Distribution",
       x= "Theoretical Normal Disribution",
       y= "Sample Distribution") + 
  geom_abline(slope=sd(Sample.sim),intercept=median(Sample.sim),size=0.8,colour="red")
```
